Using SMT engine to generate Symbolic Automata\thanks{This work was partially funded by the Associated Team FM4CPS between INRIA and ECNU, Shanghai}

### 2017-03-29 2017-Xudong-SMT/OpenPNets-SMT.tex
small typos
many todo and blue stuff

\QIN{
pNets (Parameterized Networks of Synchronized Automata) was proposed to give a behavioral specification formalism for distributed systems. It is used in the VerCors, a platform for designing and verifying distributed systems, as the intermediate language for the high-level language. This language formalizes each component of the distributed system and gives out the composition of these components. At the beginning, pNet can be used to formalize full defined systems or softwares. The open pNet is proposed to solve the problem of the uncertain components contained in the systems using "holes" as the process parameter dealing with these components. The hole acts as the placeholder for the uncertain component with a set of possible behaviors the component might conduct, presenting the "open" property. 
}

\QIN{
The pNet has developed in a series of works~\cite{HMZ:PDP15,henrio:Forte2016} and some examples has been introduced showing its ability in encoding the operators from some other algebras or the program skeletons. It is just the first step of our long-term goals. The operational semantic of the open pNet has already defined, but we only have a sketch about how to compute it and have no implementation. The operational semantic here is defined as an open automaton. The open transitions in the open automaton presenting the composition feature of the pNet is the main part of the open automaton computing and needs to satisfy several constraints. So in the computing algorithm, computing open transitions is necessary and should have some method to keep them satisfying the constraints. Then implementation of the algorithm should be finished. After getting the open automaton of the pNet, we can check the equivalence between pNets. The equivalence between pNets is "FH-bisimulation" taking the predicate of the open transitions into account each time matching such open transitions. In our long-term goals, the final target is to check the property of open systems.
}

@@

\paragraph{Context}
\TODO{
  Summary of previous works on pNets and open pNets => 1/2 page.\\
  Longterm goals:\\
  - open pNets to represent operators and program skeletons (ref to
  \cite{HMZ:PDP15})\\
  - compute semantics in term of open transitions\\
  - compute equivalence (bisimulation of open automata)\\
  - model-check properties of open systems
  }
\paragraph{Contribution}
\QIN{
This paper shows that
}

@@

\ERIC{Keep minimum defs here too be independent from FORTE}

@@

pNets can encode naturally the notion of input actions in value-passing CCS
\cite{Milner89} or of usual point-to-point message passing calculi, but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.
\ERIC{Formally, value-passing actions \emph{\`a la} CCS, denoted as ``?ch(x)'' and ``!ch(exp)'' in our CCS running example, stand respectively for $CCSact(i(ch),?x)$ and $CCSact(ch,exp)$, where $ch$ is a variable denoting a channel name, $?x$ is an input variable denoting the data argument received as input on $ch$, and $exp$ some data expression (containing no input variables). In our example we will use the usual CCS syntax. }
%For Lotos-style
%distributed synchronization, we use synchronization vectors that encode matching data
%offers on a common gate; or 
We can also use more complex action structure such as Meije-SCCS action
monoids, like in $a.b$, $a^{f(n)}$ (see \cite{deSimone85}). The expressiveness of the synchronization constructs
will depend on the action algebra.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:ccs-composed}
%  \TODO{ERIC: change for an OT of CCS example}
\ERIC{  The \texttt{CCS} pNet of Fig. \ref{schema:ccs-pnets} has 2 controllers and 2 holes. We show its full open Automaton in section \ref{section:full-result}, it has 16 open transitions. We detail 3 of them here:}

 \smallskip
 $  OT_1  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1  ~~
                       \{\}               ~~
                 [ a\neq l \land b\neq r \land a=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{00} \xrightarrow{\underline{\tau}} \ostate{10}}$

 $ OT_3  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       \{\}               ~~
                 [ a\neq l  ]              
                      }
     {\ostate{10} \xrightarrow{a} \ostate{11}} $

$  OT_{11}  = \openrule{
                          1 \xrightarrow{l}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1 ~~
                       \{ \xrightarrow{b_P}_{P} \}               ~~
                 [ a\neq l \land b\neq r \land b_P=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
  $
\end{example}
\TODO{Eric: short comments}

@@

from which we generate following three assertion:

@@

\subsection{Result of the running example}
\label{section:full-result}
% \TODO{
% Xudong: full result of the tool execution on the running example\\\\
% Open Automata of the CCS example.
\TODO{The open automaton is not the final one. The vectors are wrong,
  and you should add at least the OT name... 
}
\QIN{
We get the final result after pruning all the unsatisfiable open transitions, 2/3 of the intermediate results; the resulting open automaton is shown in Fig.~\ref{schema:ccs-result}, we give out the open automaton manually through drawing the final result open transitions generated by algorithm. A part of results we computed by hand in advance are also list below. The $ot_1$ represents the value-passing between $a.P$ and $b.Q$ and they are synchronized then give out an silent action $\tau$, process $P$ and $Q$ should not work in this case. The $ot_3$ and $ot_5$ represent another possible path to execute $a.P||b.Q$, action $a$ and $b$ are finished in sequence instead of synchronizing, they are just transmitted out as the external action. The behavior of the processes, presented as holes $P$ and $Q$ in this pNet, is uncertain, we only know the sort $Action$ for them. So the action from $P$ might be synchronized with the action $b$ that is what $ot_9$ presents or be directly transmitted to the external as the $ot_7$ shows. And we can see more cases of the interaction between these two process in the open transitions start at the state $11$, synchronizing two actions like $ot_{13}$ or directly transmitting like $ot_{15}$.
}

\QIN{
To sum up, from the pNet encoding CCS formula inputed by the user we generate the open transitions and prune the unsatisfiability automatically. The final result open transitions represent all the possible movements of the CCS example. They can exactly construct the open automaton for the given example.
}

@@

\QIN{
The next work after the generation of the open automaton of the pNet is planned to check the bisimulation of the pNets. In previous paper, we have already found the FH-bisimulaiton\cite{meiji-sccs} as a prototype. We will decide the definition of the bisimulation of the pNets and the algorithm to prove a relation between two pNets is a bisimulation. Before that work, we need to refine the result open transitions, eliminating the redundancy in the predicate of the result. The ideal result is what we showed in the bottom part of Fig.~\ref{schema:ccs-result} while there is many intermediate variables we used, such as fresh local variables of synchronization vectors, and fresh result variables of intermediate open transitions, in the algorithm for the propagation of the actions or values. So far, to eliminate the intermediate variables in the predicate, we define the range of the significant variables that we won't eliminated: the local variables of pLTS transitions, the hole behaviors and the result action at the top level. 
}

>>>>

@@

Parameterized Networks of Synchronized Automata (pNets) were proposed
to give a behavioral specification formalism for distributed
systems, synchronous, asynchronous, or heterogeneous. It is used in
VerCors, a platform for designing and 
verifying distributed systems, as the intermediate language for various
high-level languages. The high-level languages in VerCors formalize
each component of the 
distributed system and gives out the composition of these
components.
pNets provides the core low-level semantic formalism for VerCors, and
is made of a hierarchical composition of (value-passing) automatas,
called parameterized labelled transition systems (pLTS), where each
hierarchical level defines the possible synchronisation of the lower levels.
Traditionaly, pNet have been used to formalize full
defined systems or softwares. But we want also to define and reason
about incompletely defined systems, like program skeletons, operators,
or open expressions of process calculi.
The open pNet is proposed to solve the
problem of the undefined components contained in the systems using
"holes" as the process parameter dealing with these components. The
hole acts as the placeholder for the uncertain component with a set of
possible behaviors the component might conduct, presenting the "open"
property.

The pNet model was developed in a series of
works~\cite{HMZ:PDP15,henrio:Forte2016} and some examples have been
introduced showing its ability in encoding the operators from some
other algebras or  program skeletons. It is just the first step of
our long-term goals. The operational semantic of the open pNet has
already been defined, but we only had a sketch about how to compute it and
had no implementation. This operational semantic is defined as an
Open Automaton in which Open Transitions contain logical predicates
expressing the relations between the behavior of the holes, and the
global behavior of the system.

Computing the open transitions from the composition of pNet is the
main part of the open 
automaton computing and needs to satisfy the constraints expressed as
the rules of the operationnal semantics. So in
the computing algorithm, computing open transitions is necessary and
should have some method to keep them satisfying the constraints.
Computing open automata by hand is complicated and error prone. And
the algorithm published in \cite{henrio:Forte2016} was not precise
enough for an implementation. Our first goal was then to refine the
algorithm, and implement it in a tool.
In the long goal, we want to be able to check the equivalence between
pNets. The equivalence between pNets is "FH-bisimulation" taking the
predicate of the open transitions into account each time matching such
open transitions. In our long-term goals, the final target is to check
the property of open systems.

Our contribution in the article is:
\begin{itemize}
\item We have refined the open automaton generation algorithm, to make
  it implementable, and we implemented a full working prototype, within the
  VerCors platform. In the process, we had to improve slightly the
  semantics rules, and to add features in the algorithm to deal with
  management of variables and assignements.
  \item parameterisation over the action algebra operators in the
    VerCors ECORE and library
  \item interaction with the Z3 SMT solver
    \item validation on several use-cases
\end{itemize}

\TODO{I have not described the items above after th efirst one, left
  it to you...}

@@

% \ERIC{Keep minimum defs here too be independent from FORTE}

@@

pNets can encode naturally the notion of input actions in value-passing CCS
\cite{Milner89} or of usual point-to-point message passing calculi, but it also allows
for more general mechanisms, like gate negotiation in Lotos, or broadcast
communications.
Formally, value-passing actions \emph{\`a la} CCS, denoted as ``?ch(x)'' and ``!ch(exp)'' in our CCS running example, stand respectively for $CCSact(i(ch),?x)$ and $CCSact(ch,exp)$, where $ch$ is a variable denoting a channel name, $?x$ is an input variable denoting the data argument received as input on $ch$, and $exp$ some data expression (containing no input variables). In our example we will use the usual CCS syntax.
%For Lotos-style
%distributed synchronization, we use synchronization vectors that encode matching data
%offers on a common gate; or 
We can also use more complex action structure such as Meije-SCCS action
monoids, like in $a.b$, $a^{f(n)}$ (see \cite{deSimone85}). The expressiveness of the synchronization constructs
will depend on the action algebra.

@@

\begin{example}\emph{An open-transition.}
  \label{OT:ccs-composed}
%  \TODO{ERIC: change for an OT of CCS example}
The \texttt{CCS} pNet of Fig. \ref{schema:ccs-pnets} has 2 controllers and 2 holes. We show its full open Automaton in section \ref{section:full-result}, it has 16 open transitions. We detail 3 of them here:

 \smallskip
 $  OT_1  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1  ~~
                       \{\}               ~~
                 [ a\neq l \land b\neq r \land a=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{00} \xrightarrow{\underline{\tau}} \ostate{11}}$

 $ OT_3  = \openrule{
                       0 \xrightarrow{a}_{C_1} 1 ~~
                       \{\}               ~~
                 [ a\neq l  ]              
                      }
     {\ostate{10} \xrightarrow{a} \ostate{10}} $

$  OT_{11}  = \openrule{
                          1 \xrightarrow{l}_{C_1} 1 ~~
                       0 \xrightarrow{b}_{C_2} 1 ~~
                       \{ \xrightarrow{hb_P}_{P} \}               ~~
                 [ a\neq l \land b\neq r \land hb_P=?ch(v) \land b=!ch(v) ]              
                      }
     {\ostate{10} \xrightarrow{\tau} \ostate{11}}
  $
\end{example}

The first open transitions listed, $OT_1$ and $OT_3$, only involve
possible actions from the prefix operators, respectively communication
between the two prefixes, if actions $a$ and $b$ are matching emit and
receive actions on the same channel; or simple propagation of action
$a$ from the left prefix, with no special constraint other than being
different from the local $l$ action of pLTS $C_1$.

Remark the structured states in the conclusion of these rules: both
start in state $\ostate{00}$, where both $C_1$ and $C_2$ are in their
initial states. They end in different structured states, because in
$OT_1$ both pLTS chage states, meaning that the control has gone
respectively to processes $P$ and $Q$, consuming actions $A$ and $B$,
while in $OT_3$ only $C_1$ has moved.

The last open transition $OT_{11}$ starts in state $\ostate{10}$, with
control at $P$ on the left, and action $b$ on the right. So these two
can communicate, e.g. here if $b$ is an emission, and the action of hole
$P$ is the matching reception.

@@

from which we generate the following three assertions:

@@

\subsection{Result of the running example}
\label{section:full-result}
% \TODO{
% Xudong: full result of the tool execution on the running example\\\\
% Open Automata of the CCS example.

We get the final result after pruning all the unsatisfiable open
transitions, that is 2/3 of the intermediate results; the resulting open
automaton is shown in Fig.~\ref{schema:ccs-result}, we have drawn
manually the open automaton obtained as the final result generated by
algorithm.
The $ot_1$ represents the
value-passing between $a.P$ and $b.Q$ and they are synchronized then
give out an silent action $\tau$, process $P$ and $Q$ should not work
in this case. The $ot_3$ and $ot_5$ represent another possible path to
execute $a.P||b.Q$, action $a$ and $b$ are finished in sequence
instead of synchronizing, they are just transmitted out as the
external action. The behavior of the processes, presented as holes $P$
and $Q$ in this pNet, is uncertain, we only know the sort $Action$ for
them. So the action from $P$ might be synchronized with the action $b$
that is what $ot_9$ presents or be directly transmitted to the
external as the $ot_7$ shows. And we can see more cases of the
interaction between these two processes in the open transitions starting at
the state $11$, synchronizing two actions like $ot_{13}$ or directly
transmitting like $ot_{15}$.

To sum up, from the pNet encoding CCS formula inputed by the user we
generate the open transitions and prune the unsatisfiability
automatically. The final result open transitions represent all the
possible movements of the CCS example. They can exactly construct the
open automaton for the given example.

\TODO{Beware, there are incoherences between the automaton and the
  rules, e.g. OT3 goes to state 01, or 01 ?}

@@

\QIN{
The next work after the generation of the open automaton of the pNet is planned to check the bisimulation of the pNets. In previous paper, we have already found the FH-bisimulaiton\cite{deSimone85} as a prototype. We will decide the definition of the bisimulation of the pNets and the algorithm to prove a relation between two pNets is a bisimulation. Before that work, we need to refine the result open transitions, eliminating the redundancy in the predicate of the result. The ideal result is what we showed in the bottom part of Fig.~\ref{schema:ccs-result} while there is many intermediate variables we used, such as fresh local variables of synchronization vectors, and fresh result variables of intermediate open transitions, in the algorithm for the propagation of the actions or values. So far, to eliminate the intermediate variables in the predicate, we define the range of the significant variables that we won't eliminated: the local variables of pLTS transitions, the hole behaviors and the result action at the top level. 
}
